Converging High-Throughput and High-Performance Computing: A Case Study

### 2017-03-29 Experiments.tex
Pass on NGE experiments and scattered space savings

In this section we present experiments that aim to determine the performances of the NGE and to show how it provides a minimal overhead while introducing new functionalities.

Experiments consist in running AthenaMP instances by using NGE pilots where each AthenaMP simulates a pre-determined number of events taken from ATLAS workload.	
We present three groups of experiments in which we test the NGE for weak scalability, weak scalability with multiple generation and strong scalability.

We collected data about the execution time of the pilots and the AthenaMP that have been executed within them by considering all stages of the execution.
Experiments have been performed out of production by using NGE's pilots on TITAN's batch queue. 
Because of TITAN's batch queue, the turnaround time of each run of our experiments is dominated by the time spent on TITAN's queue. Since we are interested only in the performances of the NGE, we removed queue time from our statistics.

All the experiments have been performed by setting AthenaMP in such a way that it uses all the 16 cores present on TITAN's nodes.

@@

The experiment consists in running as many AthenaMP instances (also referred as tasks from now-on) as the number of nodes controlled by the pilot. Each AthenaMP have been assigned with 100 events whose simulation requires  ~$70$ minutes.

Tasks do not experience queue within the pilot since there is one node for each AthenaMP instance. Therefore, delays are consequence of only three factors: i) the bootstrapping of the pilot on the nodes; ii) the manager, as defined in Section \ref{sec:arch}, that has to dispatch tasks to the pilot; iii) the time that the agent requires to spawn all the tasks on the nodes.

We tested different pilot sizes, i.e. : 250, 500, 1000 and 2000. For all of them the walltime was 2 hours.

Figure \ref{fig:weakScal1a} depicts the average pilot duration, the average execution time of AthenaMP and pilot's overhead as function of the pilot size.

\begin{figure}[!htb]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
    \caption{Weak scalability: average pilot duration, average duration of a single AthenaMP execution and pilot's overhead for different pilot sizes (250, 500, 1000 and 2000 nodes). Note that the overhead refers to the axis on the right and it is expressed in tenths of a second.}
\label{fig:weakScal1a}
\end{figure}

Firstly, we can observe that, despite some fluctuations due to external factors\footnote{such as the Titan's shared filesystem and the shared database used by the NGE}, the average execution time of AthenaMP oscillates between 4200 and 4800 seconds. Secondly, we can observe that in all the cases the gap between AthenaMP execution times and the pilot durations is minimal although it slightly increases with the pilot size.
By observing NGE's overhead, we can notice the overhead does not grow linearly with the number of units.

@@

\subsubsection{Weak scalability with multiple generation }
This experiment is similar to the one presented above but in this case we want to test also the impact of submitting new AthenaMP instances in place of those that end. The reason behind this experiment is to stress pilot's components by pushing new tasks while others are ending their execution.

In order to perform the experiments in less than two hours per run, we decided to send for execution a number of AthenaMP instances equal to five times the number of nodes and we reduced  the number of events that are simulated by each Athena-MP to sixteen\footnote{Note that the number of events has been chosen in such a way that all the cores of a node run one event.}. This allows us to complete a single AthenaMP in ~20 minutes only.

We tested different pilot sizes, i.e. : 256, 512, 1024 and 2048. For all of them the walltime was 3 hours.

Figure \ref{fig:weakScal2a} depicts the average pilot duration, the average execution time of five sequential instances of AthenaMP and the corresponding overhead.  
We can observe that the difference between the two durations is more marked than in the previous experiments. Despite this we can notice that the growth of the overhead is still less than linear.

\begin{figure}[!htb]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak2.pdf}
    \caption{Weak scalability with multiple generations: average pilot duration, average duration of five sequential AthenaMP executions and pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes). Note that the overhead refers to the axis on the right and it is expressed in tenths of a second.}
\label{fig:weakScal2a}
\end{figure}
%\begin{figure}[!htb]
%        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakOver2.pdf}
%    \caption{Weak scalability with multiple generations: average overhead by running AthenaMP on 256, 512, 1024 and 2048 nodes.}
%\label{fig:weakScal2b}
%\end{figure}
\subsubsection{Strong scalability}
The last experiments deals with strong scalability. Therefore, we run the same amount of tasks for different pilot sizes. 
We used a number of AthenaMP instances equal to 2048 and tested pilots with 256, 512, 1024 and 2048 nodes. Thus, the number of sequential AthenaMP instances is equal to eight for the smallest pilot and is null for the largest.  
Each AthenaMP instance simulates sixteen events as the previous experiment.

Figure \ref{fig:strongScala} depicts the average pilot duration and  the average execution time of possibly sequential AthenaMP instances.  We can notice that the difference between the pilot duration and the AthenaMP execution times is almost constant for all the pilot sizes although the overall duration of the pilot decreases linearly with the pilot size.

\begin{figure}[!htb]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/strong.pdf}
    \caption{Strong scalability:  average pilot duration, average duration of sequential AthenaMP executions and pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes). Note that the overhead refers to the axis on the right and it is expressed in tenths of a second.}
\label{fig:strongScala}
\end{figure}

>>>>

In this section we present experiments % that aim
to determine the performances of the NGE and to show how it provides a minimal
overhead while introducing new functionalities.

Experiments % consist in running
run instances of AthenaMP by using NGE pilots, % where each AthenaMP simulates
simulating a pre-determined number of events in the ATLAS detector.\mtnote{are
they 100 events?}
% taken from ATLAS workload.
We present three groups of experiments in which we test the NGE for weak
scalability, weak scalability with multiple generation, and strong scalability.

We % collected data about
measured the execution time of the pilots and of the AthenaMP % that have been
executed within them, collecting timestamps at % by considering
all stages of the execution. Experiments have been performed % out of production
by % using
submitting NGE's pilots % on
to TITAN's batch queue. Because of TITAN's % batch queue
policies, the turnaround time of each run of our experiments is dominated by
% the time spent on TITAN's queue
queue time. Since we are interested only in the performances of the NGE, we
removed queue time from our statistics.

All the experiments have been performed by % setting
configuring AthenaMP
% in such a way that it uses
to use all the 16 cores % present on
of TITAN's worker nodes.

@@

% The
In this experiment % consists in running
we run as many AthenaMP instances (also referred as tasks from now-on) as the number of nodes controlled by the pilot. Each AthenaMP
% have been assigned with 100 events whose simulation requires
simulates 100 events, requiring ~$70$ minutes.

Tasks do not % experience
wait within the NGE Agent's queue % within the pilot
since % there is
one node %for
is available to each AthenaMP instance. % Therefore, d
Delays in tasks execution are consequence of % only
three other factors: (i) the bootstrapping of the pilot on the nodes; (ii) the
UnitManager, as defined in Section~\ref{sec:arch}, that has to dispatch tasks to
the agent; and (iii) the time that the agent requires to spawn all the tasks on
the nodes.

We tested % different
pilots % sizes, i.e. :
with 250, 500, 1000 and 2000 worker nodes and
%. For all of them, the walltime was
2 hours walltime. Figure~\ref{fig:weakScal1a} depicts the average pilot
duration, the average execution time of AthenaMP, and the pilot's overhead as
function of the pilot size.

\begin{figure}[!htb]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak1.pdf}
    \caption{Weak scalability: average pilot duration, average duration of a
    single AthenaMP execution, and pilot's overhead for different pilot sizes
    (250, 500, 1000 and 2000 worker nodes). Note: the overhead refers to
    the axis on the right and it is expressed in tenths of a second.}
\label{fig:weakScal1a}
\end{figure}

% Firstly, w
We % can
observe that, despite some fluctuations due to external factors
% \footnote{such as the
(e.g., Titan's shared filesystem and the shared database used by the NGE), the
average execution time of AthenaMP % oscillates
ranges between 4200 and 4800 seconds. % Secondly, w
We % can
also observe that in all the cases the gap between AthenaMP execution times and
the pilot durations is minimal, although it slightly increases with the pilot
size. % By observing NGE's overhead, w
We % can
notice that NGE's overhead does not grow linearly with the number of units.

@@

\subsubsection{Weak scalability with multiple generation }

This experiment is similar to the one presented above but, in this case, we want
to test also the impact of submitting multiple generations of % new
AthenaMP to the same pilot.\mtnote{we used `generations' at the beginning of 6,
I am assuming the reader understands the term by now. Please change if you
disagree.}
% instances in place of those that end.
% The reason behind this experiment is to
In this way, we stress the pilot's components, % by pushing new tasks while others are ending their execution
as new tasks are scheduled for execution on the Agent while other tasks are still running.

In order to perform the experiments in less than 2 hours per run\mtnote{we do
we need to keep it below 2 hours?}, we
% decided to send for execution a number of AthenaMP instances equal to
used five times AthenaMP instances than the number of worker nodes. % and w
We also reduced  the number of events
% that are simulated
simulated by each AthenaMP to sixteen.
% \footnote{Note that the number of events has been chosen in such a way
% that all the cores of a node run one event.}
In this way, every core of each worker node was used to simulate one event,
%. This allows us to complete
reducing the running time of each AthenaMP to ~$20$ minutes. % only.

We tested
% different
pilots
% sizes, i.e. :
with 256, 512, 1024 and 2048 worker nodes and
%. For all of them, the walltime was
3\mtnote{2?} hours walltime. Figure~\ref{fig:weakScal2a} depicts the average
pilot duration, the average execution time of five sequential
instances\mtnote{generations?} of AthenaMP, and the corresponding overhead. We
% can
observe that the difference between the two durations is more marked than in the
previous experiments \mtnote{mostly due to the increased overhead?}. Despite
this, we
% can
notice that the growth of the overhead is still less than linear.

\begin{figure}[!htb]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/weak2.pdf}
    \caption{Weak scalability with multiple generations: average pilot
    duration, average duration of five sequential AthenaMP executions, and
    pilot's overhead for different pilot sizes (256, 512, 1024 and 2048 nodes).
    Note that the overhead refers to the axis on the right and it is expressed
    in tenths of a second.}
\label{fig:weakScal2a}
\end{figure}

%\begin{figure}[!htb]
%        \includegraphics[width=0.4\textwidth]{./figures/NGE/weakOver2.pdf}
%    \caption{Weak scalability with multiple generations: average overhead by running AthenaMP on 256, 512, 1024 and 2048 nodes.}
%\label{fig:weakScal2b}
%\end{figure}

\subsubsection{Strong scalability}

The last experiments % deals with
study strong scalability. % Therefore, we
by running the same amount of tasks for different pilot sizes. We used
% a number of AthenaMP instances equal to
2048 AthenaMP instances and % tested
pilots with 256, 512, 1024 and 2048 nodes. Thus, the number of sequential
AthenaMP instances is equal to 8 for the smallest pilot and is null for the
largest\mtnote{Not sure I understand this sentence}. Each AthenaMP instance
simulates sixteen events as in the previous experiment.

Figure \ref{fig:strongScala} % depicts
shows the average pilot duration and the average execution time of possibly
sequential AthenaMP instances.  We % can
notice that the difference between the pilot duration and the AthenaMP execution
times is almost constant for all the pilot sizes, although the overall duration
of the pilot decreases linearly with the pilot size.

\begin{figure}[!htb]
        \includegraphics[height=4.5cm,width=\columnwidth]{./figures/NGE/strong.pdf}
    \caption{Strong scalability:  average pilot duration, average duration of
    sequential AthenaMP executions, and pilot's overhead for different pilot
    sizes (256, 512, 1024 and 2048 nodes). Note that the overhead refers to the
    axis on the right and it is expressed in tenths of a second.}
\label{fig:strongScala}
\end{figure}

### 2017-03-29 PandaOverview.tex
Pass on NGE experiments and scattered space savings

PanDA is a Workload Management System (WMS) designed to support the execution of
distributed workloads and workflows via pilots. WMS is middleware for
discovering and selecting resources, submitting tasks of workloads and
workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
abstraction that enables multi-level scheduling by decoupling resource
acquisition from tasks scheduling~\cite{turilli2015comprehensive}. Once
implemented, a pilot is scheduled on a site (e.g., by submitting a job or
requesting a virtual machine) and, once active, tasks are scheduled to the
pilot, not to the site's scheduler.

Pilot-enabled WMS are particularly relevant for LHC experiments, where millions
of tasks are executed across multiple sites every month, analyzing and producing
petabytes of data. The design of PanDA WMS started in 2005 and its
implementation went into production for the LHC Run 1 on 2009. PanDA was then
extended with new subsystems to be deployed on Run 2, on 2015. We summarize
PanDA's design, architecture, and execution process, focussing on those
subsystems and components that are most relevant to the the integration of PanDA
with supercomputers.

>>>>

PanDA is a Workload Management System (WMS) designed to support the execution of
distributed workloads and workflows via pilots. WMS is middleware for
discovering and selecting resources, submitting tasks of workloads and
workflows, and monitoring their execution~\cite{marco2009glite}. Pilot is an
abstraction that enables multi-level scheduling by decoupling resource
acquisition from tasks scheduling~\cite{turilli2015comprehensive}. When
implemented, a pilot is scheduled on a site and, once active, tasks are
scheduled to the pilot, not to the site's scheduler.

Pilot-enabled WMS enable high throughput of tasks execution while supporting
interoperability across multiple sites. This is particularly relevant for LHC
experiments, where millions of tasks are executed across multiple sites every
month, analyzing and producing petabytes of data. The design of PanDA WMS
started in 2005 and its implementation went into production for the LHC Run 1 on
2009. PanDA was then extended with new subsystems to be deployed on Run 2, on
2015.

% We summarize PanDA's design, architecture, and execution process, focussing on
% those subsystems and components that are most relevant to the the integration
% of PanDA with supercomputers.
